---
title: "Analyse"
author: "Adrian Stämpfli"
date: "13 4 2021"
output:
  html_document:
    df_print: paged
---

```{r, include = FALSE}
library(tidyverse)
library(kml)
library(GGally)
library(plm)
library(stargazer)
library(BBmisc)
library(REEMtree)
library(rattle)
library(knitr)

load("../data/df_long.R")



colnames(df_long) <- c("id","year","depression","ermuedung","stress_arbeit","rueckenschmerzen","arbeit_zufriedenheit_aufgaben","art_arbeitszeit","arbeit_einbezug_entscheidungen","arbeitskontrolle_personen","arbeit_qualifikation","arbeitsstunden_woche","wochenend_arbeit","nachtarbeit","taegl_pendeln_min","arbeit_intensitaet","zufriedenheit_arbeitsatmosphaere","arbeit_laerm_schmutz","arbeit_ermuedende_koerperha","hausarbeit_stunden_woche","beeintraechtigung_arbeit_privat","abschalten_nach_arbeit","einschraenkung_weg_ges_zustand","tage_gesunheits_prob","chronische_krankheit","ausbildung","partnerschaft","tod_person","person_haushalt","migrationshintergrund","geschlecht","alter","status","occupa","ch_nationalitaet","haushaltsaequivalenzeinkommen","kinder_betreuung")




sample <- df_long %>%
  drop_na(alter) %>%
  filter(alter >= 15) %>%
  filter(alter <= 65) %>%
  filter(occupa %in% c(1,2,3,5,6))



sample <- as.data.frame(sapply(sample, as.numeric))
```


## KML

Clustersize 2-8, Redrawing fixed to 5 (Redrawing does not really makes a differnce in the output).


```{r, echo = FALSE}
### Run KML
df_kml = sample[1:3]

for (cluster_number in c(5)) {
    for (redrawing_number in c(5)) {
      kml_cluster_data <- reshape(df_kml, idvar = "id", timevar = "year", direction = "wide")
      cluster = clusterLongData(as.matrix(kml_cluster_data[c(2:16)]))

      start_time <- Sys.time()
      kml(cluster, nbClusters = cluster_number, nbRedrawing = redrawing_number, toPlot='none')
      end_time <- Sys.time()
      runtime = end_time - start_time
      plot(cluster,cluster_number)
      print("KML Algortihm")
      print(runtime)
      print(paste0("Cluster number = ", cluster_number, ". Redrawing number = ", redrawing_number))
    }
  }

```


\pagebreak
## KML Cluster = 4

For further processing we use KML with a clustersize of 4.


```{r, echo = FALSE}
cluster_size = 4
kml_cluster_data <- reshape(df_kml, idvar = "id", timevar = "year", direction = "wide")
cluster = clusterLongData(as.matrix(kml_cluster_data[c(2:16)]))
Shape_kml = kml(cluster, nbClusters = cluster_size, nbRedrawing = 5, toPlot='none')

plot(cluster,cluster_size)
```


```{r, include = FALSE}
### code to add clusters to the original data
id_not_na = as.numeric(str_extract(cluster@idFewNA, "\\-*\\d+\\.*\\d*"))
unique_id = unique(df_kml$id)
labels = as.data.frame(cluster@c4[[5]]@clusters)


real_id_not_na <- list()
for(i in 1:length(id_not_na))
{ 

  real_id_not_na[[i]] <- unique_id[id_not_na[i]]

}
real_id_not_na <- do.call(rbind.data.frame, real_id_not_na)
df_labeled <- cbind(real_id_not_na, labels)
colnames(df_labeled) <- c("id","cluster")


df_clustered = right_join(sample, df_labeled, by = "id")

```

Added the KML clusters to the original data frame.

```{r}
kable(head(df_clustered[c(1,2,3,38)]))
```

\pagebreak
## Descriptive analysis of cluster ~ variable

First, standardize all continous variables.
Then reranged all variables to 0-1.

```{r, include=FALSE}
### Scale all Variables which are not categorical
df_clustered_not_scaled = df_clustered
df_clustered = df_clustered %>% mutate_at(c('taegl_pendeln_min',"arbeitsstunden_woche","hausarbeit_stunden_woche","tage_gesunheits_prob","person_haushalt","haushaltsaequivalenzeinkommen"), ~(scale(.) %>% as.vector))
### Rerange all variables to 0-1
df_clustered = normalize(df_clustered, method = "range", range = c(0, 1))

```


```{r, echo = FALSE, warning = FALSE, message = FALSE}
# 4,7,11,14,15,16,19,20,21,22,23,24,26,32,36
df_plot = df_clustered[c(4,7,11,38)]
p <- ggpairs(df_plot, aes(color = cluster, alpha = 0.3)) + theme_bw()

for(i in 1:p$nrow) {
  for(j in 1:p$ncol){
    p[i,j] <- p[i,j] +
      scale_fill_manual(values=c("#1B9E77", "#D95F02", "#7570B3", "#E7298A")) +
      scale_color_manual(values=c("#1B9E77", "#D95F02", "#7570B3", "#E7298A"))
  }
}
p

df_plot = df_clustered[c(12,16,17,38)]
p <- ggpairs(df_plot, aes(color = cluster, alpha = 0.3)) + theme_bw()

for(i in 1:p$nrow) {
  for(j in 1:p$ncol){
    p[i,j] <- p[i,j] +
      scale_fill_manual(values=c("#1B9E77", "#D95F02", "#7570B3", "#E7298A")) +
      scale_color_manual(values=c("#1B9E77", "#D95F02", "#7570B3", "#E7298A"))
  }
}
p

df_plot = df_clustered[c(20,21,22,38)]
p <- ggpairs(df_plot, aes(color = cluster, alpha = 0.3)) + theme_bw()

for(i in 1:p$nrow) {
  for(j in 1:p$ncol){
    p[i,j] <- p[i,j] +
      scale_fill_manual(values=c("#1B9E77", "#D95F02", "#7570B3", "#E7298A")) +
      scale_color_manual(values=c("#1B9E77", "#D95F02", "#7570B3", "#E7298A"))
  }
}
p

df_plot = df_clustered[c(23,26,32,38)]
p <- ggpairs(df_plot, aes(color = cluster, alpha = 0.3)) + theme_bw()

for(i in 1:p$nrow) {
  for(j in 1:p$ncol){
    p[i,j] <- p[i,j] +
      scale_fill_manual(values=c("#1B9E77", "#D95F02", "#7570B3", "#E7298A")) +
      scale_color_manual(values=c("#1B9E77", "#D95F02", "#7570B3", "#E7298A"))
  }
}
p
```

\pagebreak
## Regressionsanalysis




### Pooled OLS Variableset 1
```{r, include=FALSE}
df_clustered$cluster = as.numeric(df_clustered$cluster)
```


```{r, echo=FALSE}
reg1 = plm(depression ~ ausbildung + geschlecht + ch_nationalitaet, data = df_clustered, index = c("id","year"), model="pooling")
stargazer(reg1,   
          title="Pooled OLS", type="text", 
          column.labels=c("Pooled OLS VS1"), 
          df=FALSE, digits=4)
```

```{r, echo=FALSE}
kable(sort(reg1$coefficients, decreasing = TRUE))
```

\pagebreak
### Pooled OLS Variableset 2

```{r, echo=FALSE}
reg2 = plm(depression ~ stress_arbeit + arbeit_zufriedenheit_aufgaben + art_arbeitszeit + arbeit_einbezug_entscheidungen + arbeitskontrolle_personen + arbeit_qualifikation + arbeitsstunden_woche + wochenend_arbeit + nachtarbeit + taegl_pendeln_min + arbeit_intensitaet + zufriedenheit_arbeitsatmosphaere + arbeit_laerm_schmutz + arbeit_ermuedende_koerperha + einschraenkung_weg_ges_zustand + tage_gesunheits_prob + chronische_krankheit + ausbildung + partnerschaft + tod_person + migrationshintergrund + geschlecht + alter + ch_nationalitaet + haushaltsaequivalenzeinkommen, data = df_clustered, index = c("id","year"), model="pooling")
stargazer(reg2,   
          title="Pooled OLS", type="text", 
          column.labels=c("Pooled OLS VS2"), 
          df=FALSE, digits=4)



```

```{r, echo=FALSE}
kable(sort(reg2$coefficients, decreasing = TRUE))
```


\pagebreak
### Fixed Effects

```{r, echo=FALSE}
reg3 = plm(depression ~ arbeit_zufriedenheit_aufgaben + art_arbeitszeit + arbeit_einbezug_entscheidungen + arbeitskontrolle_personen + arbeit_qualifikation + arbeitsstunden_woche + wochenend_arbeit + nachtarbeit + taegl_pendeln_min + arbeit_intensitaet + zufriedenheit_arbeitsatmosphaere + arbeit_laerm_schmutz + arbeit_ermuedende_koerperha, data = df_clustered, index = c("id","year"), model="within")
stargazer(reg3,   
          title="Fixed Effects", type="text", 
          column.labels=c("Fixed Effects"), 
          df=FALSE, digits=4)
```

```{r, echo=FALSE}
kable(sort(reg3$coefficients, decreasing = TRUE))
```


\pagebreak
## Regressionsanalysis per Cluster
### Pooled OLS Variableset 1

```{r, echo=FALSE}
regressions = list()
for(i in 1:length(unique(df_clustered$cluster))){
  df_per_cluster = df_clustered %>% filter(cluster == i)
  regressions[[i]] = plm(depression ~ ausbildung + geschlecht + ch_nationalitaet, data = df_per_cluster, index = c("id","year"), model="pooling")
}
  
stargazer(regressions[[1]], regressions[[2]], regressions[[3]], regressions[[4]],   
          title="Pooled OLS VS1", type="text", 
          column.labels=c("Cluster 1", "Cluster 2", "Cluster 3", "Cluster 4"), 
          df=FALSE, digits=4)
```

\pagebreak
### Pooled OLS Variableset 2

```{r, echo=FALSE}
regressions = list()
for(i in 1:length(unique(df_clustered$cluster))){
  df_per_cluster = df_clustered %>% filter(cluster == i)
  regressions[[i]] = plm(depression ~ stress_arbeit + arbeit_zufriedenheit_aufgaben + art_arbeitszeit + arbeit_einbezug_entscheidungen + arbeitskontrolle_personen + arbeit_qualifikation + arbeitsstunden_woche + wochenend_arbeit + nachtarbeit + taegl_pendeln_min + arbeit_intensitaet + zufriedenheit_arbeitsatmosphaere + arbeit_laerm_schmutz + arbeit_ermuedende_koerperha + einschraenkung_weg_ges_zustand + tage_gesunheits_prob + chronische_krankheit + ausbildung + partnerschaft + tod_person + migrationshintergrund + geschlecht + alter + ch_nationalitaet + haushaltsaequivalenzeinkommen, data = df_per_cluster, index = c("id","year"), model="pooling")
}

stargazer(regressions[[1]], regressions[[2]], regressions[[3]], regressions[[4]],   
          title="Pooled OLS VS2", type="text", 
          column.labels=c("Cluster 1", "Cluster 2", "Cluster 3", "Cluster 4"), 
          df=FALSE, digits=4)
```

\pagebreak
### Fixed Effects

```{r, echo=FALSE}
regressions = list()
for(i in 1:length(unique(df_clustered$cluster))){
  df_per_cluster = df_clustered %>% filter(cluster == i)
  regressions[[i]] = plm(depression ~ arbeit_zufriedenheit_aufgaben + art_arbeitszeit + arbeit_einbezug_entscheidungen + arbeitskontrolle_personen + arbeit_qualifikation + arbeitsstunden_woche + wochenend_arbeit + nachtarbeit + taegl_pendeln_min + arbeit_intensitaet + zufriedenheit_arbeitsatmosphaere + arbeit_laerm_schmutz + arbeit_ermuedende_koerperha, data = df_per_cluster, index = c("id","year"), model="within")
}

stargazer(regressions[[1]], regressions[[2]], regressions[[3]], regressions[[4]],   
          title="Fixed Effects", type="text", 
          column.labels=c("Cluster 1", "Cluster 2", "Cluster 3", "Cluster 4"), 
          df=FALSE, digits=4)
```

\pagebreak
## Regression Tree
Regression Trees with Random Effects for Longitudinal (Panel) Data

https://cran.r-project.org/web/packages/REEMtree/REEMtree.pdf

Example

https://quantdev.ssri.psu.edu/tutorials/apa-ati-intensive-longitudinal-data-session-t-random-effect-tree-models

### Regression tree variableset 1



```{r, echo=FALSE, warning=FALSE}
set.seed(1234)
tree = REEMtree(depression ~  ausbildung + geschlecht + ch_nationalitaet, data = df_clustered_not_scaled %>% drop_na(), random=~1|id)
fancyRpartPlot(tree(tree),
               digits = 2,
               sub="",
               main="")
```

### Regression tree all variables

```{r, echo=FALSE, warning=FALSE}
set.seed(1234)
tree = REEMtree(depression ~ stress_arbeit + rueckenschmerzen + arbeit_zufriedenheit_aufgaben + art_arbeitszeit +               arbeit_einbezug_entscheidungen + arbeitskontrolle_personen + arbeit_qualifikation + arbeitsstunden_woche + wochenend_arbeit + nachtarbeit +         taegl_pendeln_min + arbeit_intensitaet + zufriedenheit_arbeitsatmosphaere + arbeit_laerm_schmutz + arbeit_ermuedende_koerperha +    hausarbeit_stunden_woche + beeintraechtigung_arbeit_privat +abschalten_nach_arbeit + einschraenkung_weg_ges_zustand + tage_gesunheits_prob + chronische_krankheit + ausbildung + partnerschaft + tod_person + person_haushalt  + migrationshintergrund + geschlecht + alter + status + ch_nationalitaet + haushaltsaequivalenzeinkommen + kinder_betreuung, data = df_clustered_not_scaled %>% drop_na(), random=~1|id)
fancyRpartPlot(tree(tree),
               digits = 2,
               sub="",
               main="")
```

```{r, echo=FALSE, warning=FALSE}
df_clustered_not_scaled %>% ggplot(aes(cluster, einschraenkung_weg_ges_zustand, fill=cluster)) +
  geom_boxplot()
```

```{r, echo=FALSE, warning=FALSE}
df_clustered_not_scaled %>% ggplot(aes(cluster, abschalten_nach_arbeit, fill=cluster)) +
  geom_boxplot()
```

```{r}

```



\pagebreak
## KML Cluster = 5

For further processing we use KML with a clustersize of 4.


```{r, echo = FALSE}
cluster_size = 5
kml_cluster_data <- reshape(df_kml, idvar = "id", timevar = "year", direction = "wide")
cluster = clusterLongData(as.matrix(kml_cluster_data[c(2:16)]))
Shape_kml = kml(cluster, nbClusters = cluster_size, nbRedrawing = 5, toPlot='none')

plot(cluster,cluster_size)
```


```{r, include = FALSE}
### code to add clusters to the original data
id_not_na = as.numeric(str_extract(cluster@idFewNA, "\\-*\\d+\\.*\\d*"))
unique_id = unique(df_kml$id)
labels = as.data.frame(cluster@c5[[5]]@clusters)


real_id_not_na <- list()
for(i in 1:length(id_not_na))
{ 

  real_id_not_na[[i]] <- unique_id[id_not_na[i]]

}
real_id_not_na <- do.call(rbind.data.frame, real_id_not_na)
df_labeled <- cbind(real_id_not_na, labels)
colnames(df_labeled) <- c("id","cluster")


df_clustered = right_join(sample, df_labeled, by = "id")

```

Added the KML clusters to the original data frame.

```{r}
kable(head(df_clustered[c(1,2,3,38)]))
```

\pagebreak
## Descriptive analysis of cluster ~ variable

First, standardize all continous variables.
Then reranged all variables to 0-1.

```{r, include=FALSE}
### Scale all Variables which are not categorical
df_clustered_not_scaled = df_clustered
df_clustered = df_clustered %>% mutate_at(c('taegl_pendeln_min',"arbeitsstunden_woche","hausarbeit_stunden_woche","tage_gesunheits_prob","person_haushalt","haushaltsaequivalenzeinkommen"), ~(scale(.) %>% as.vector))
### Rerange all variables to 0-1
df_clustered = normalize(df_clustered, method = "range", range = c(0, 1))

```


```{r, echo = FALSE, warning = FALSE, message = FALSE}
# 4,7,11,14,15,16,19,20,21,22,23,24,26,32,36
df_plot = df_clustered[c(4,7,11,38)]
p <- ggpairs(df_plot, aes(color = cluster, alpha = 0.3)) + theme_bw()

for(i in 1:p$nrow) {
  for(j in 1:p$ncol){
    p[i,j] <- p[i,j] +
      scale_fill_manual(values=c("#1B9E77", "#D95F02", "#7570B3", "#E7298A", "#66A61E")) +
      scale_color_manual(values=c("#1B9E77", "#D95F02", "#7570B3", "#E7298A", "#66A61E"))
  }
}
p

df_plot = df_clustered[c(12,16,17,38)]
p <- ggpairs(df_plot, aes(color = cluster, alpha = 0.3)) + theme_bw()

for(i in 1:p$nrow) {
  for(j in 1:p$ncol){
    p[i,j] <- p[i,j] +
      scale_fill_manual(values=c("#1B9E77", "#D95F02", "#7570B3", "#E7298A", "#66A61E")) +
      scale_color_manual(values=c("#1B9E77", "#D95F02", "#7570B3", "#E7298A", "#66A61E"))
  }
}
p

df_plot = df_clustered[c(20,21,22,38)]
p <- ggpairs(df_plot, aes(color = cluster, alpha = 0.3)) + theme_bw()

for(i in 1:p$nrow) {
  for(j in 1:p$ncol){
    p[i,j] <- p[i,j] +
      scale_fill_manual(values=c("#1B9E77", "#D95F02", "#7570B3", "#E7298A", "#66A61E")) +
      scale_color_manual(values=c("#1B9E77", "#D95F02", "#7570B3", "#E7298A", "#66A61E"))
  }
}
p

df_plot = df_clustered[c(23,26,32,38)]
p <- ggpairs(df_plot, aes(color = cluster, alpha = 0.3)) + theme_bw()

for(i in 1:p$nrow) {
  for(j in 1:p$ncol){
    p[i,j] <- p[i,j] +
      scale_fill_manual(values=c("#1B9E77", "#D95F02", "#7570B3", "#E7298A","#66A61E")) +
      scale_color_manual(values=c("#1B9E77", "#D95F02", "#7570B3", "#E7298A","#66A61E"))
  }
}
p
```

\pagebreak
## Regressionsanalysis





```{r, include=FALSE}
df_clustered$cluster = as.numeric(df_clustered$cluster)
```


\pagebreak
## Regressionsanalysis per Cluster
### Pooled OLS Variableset 1

```{r, echo=FALSE}
regressions = list()
for(i in 1:length(unique(df_clustered$cluster))){
  df_per_cluster = df_clustered %>% filter(cluster == i)
  regressions[[i]] = plm(depression ~ ausbildung + geschlecht + ch_nationalitaet, data = df_per_cluster, index = c("id","year"), model="pooling")
}
  
stargazer(regressions[[1]], regressions[[2]], regressions[[3]], regressions[[4]], regressions[[5]],   
          title="Pooled OLS VS1", type="text", 
          column.labels=c("Cluster 1", "Cluster 2", "Cluster 3", "Cluster 4", "Cluster 5"), 
          df=FALSE, digits=4)
```

\pagebreak
### Pooled OLS Variableset 2

```{r, echo=FALSE}
regressions = list()
for(i in 1:length(unique(df_clustered$cluster))){
  df_per_cluster = df_clustered %>% filter(cluster == i)
  regressions[[i]] = plm(depression ~ stress_arbeit + arbeit_zufriedenheit_aufgaben + art_arbeitszeit + arbeit_einbezug_entscheidungen + arbeitskontrolle_personen + arbeit_qualifikation + arbeitsstunden_woche + wochenend_arbeit + nachtarbeit + taegl_pendeln_min + arbeit_intensitaet + zufriedenheit_arbeitsatmosphaere + arbeit_laerm_schmutz + arbeit_ermuedende_koerperha + einschraenkung_weg_ges_zustand + tage_gesunheits_prob + chronische_krankheit + ausbildung + partnerschaft + tod_person + migrationshintergrund + geschlecht + alter + ch_nationalitaet + haushaltsaequivalenzeinkommen, data = df_per_cluster, index = c("id","year"), model="pooling")
}

stargazer(regressions[[1]], regressions[[2]], regressions[[3]], regressions[[4]], regressions[[5]],   
          title="Pooled OLS VS1", type="text", 
          column.labels=c("Cluster 1", "Cluster 2", "Cluster 3", "Cluster 4", "Cluster 5"), 
          df=FALSE, digits=4)
```

\pagebreak
### Fixed Effects

```{r, echo=FALSE}
regressions = list()
for(i in 1:length(unique(df_clustered$cluster))){
  df_per_cluster = df_clustered %>% filter(cluster == i)
  regressions[[i]] = plm(depression ~ arbeit_zufriedenheit_aufgaben + art_arbeitszeit + arbeit_einbezug_entscheidungen + arbeitskontrolle_personen + arbeit_qualifikation + arbeitsstunden_woche + wochenend_arbeit + nachtarbeit + taegl_pendeln_min + arbeit_intensitaet + zufriedenheit_arbeitsatmosphaere + arbeit_laerm_schmutz + arbeit_ermuedende_koerperha, data = df_per_cluster, index = c("id","year"), model="within")
}

stargazer(regressions[[1]], regressions[[2]], regressions[[3]], regressions[[4]], regressions[[5]],   
          title="Pooled OLS VS1", type="text", 
          column.labels=c("Cluster 1", "Cluster 2", "Cluster 3", "Cluster 4", "Cluster 5"), 
          df=FALSE, digits=4)
```





```{r}

```


